Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-05 paper.tex
citation changes

The AI and machine learning literature contains two broad takes on this problem.
The first take is that the agent should come up with a better representation of the space of solutions,
for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
Our contribution is a general algorithm for fusing these two takes on the problem:
we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
alongside a bottom-up discriminative model that regresses from problems to solutions.
We evaluate our algorithm on four domains:
building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
Because we want a domain-general
approach,
we model solutions to tasks as
programs,
which frames our problem as one of program induction.

@@

TODO: talk about existing works as special cases of, and points of inspiration for, \system:
\begin{itemize}
  \item Maximum-likelihood estimation of a PCFG \citep{DBLP:conf/icml/LiangJK10} with
    production probabilities $\theta$, a fixed the DSL $\mathcal{D}$ and no neural network
    $q$;
    % although some of the Liang paper actually learns subexpressions and therefore DSL
  \item The EC algorithm \citep{Dechter:2013:BLV:2540128.2540316} has no recognition model
    $q$;
    % because this is a primary inspiration, might be worth further mention:
    % different program representation (routing vs. substitution; simple vs polymorphic types);
    % different grammar induction (sequitur vs. fragment grammar)
  \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
    DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
    network $q$.
    % although RobustFill has very different enumeration
\end{itemize}

@@

\textbf{Neural recognition model: tractably maxing $J$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
frontier $\mathcal{F}_x$ by searching for programs maximizing
$q(p|x)$.
The network $q$ exploits the structure of the DSL $\mathcal{D}$:
rather than directly predicting a distribution over $p$ conditioned on $x$,
it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
This approach implements an amortized
inference scheme~\cite{ritchie2016deep} for the generative model in
Fig.~\ref{graphicalModel}.

>>>>

The AI and machine learning literature contains two broad takes on this problem.
The first take is that the agent should come up with a better representation of the space of solutions,
for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
Our contribution is a general algorithm for fusing these two takes on the problem:
we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
alongside a bottom-up discriminative model that regresses from problems to solutions.
We evaluate our algorithm on four domains:
building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
%Because we want a domain-general
%approach,
We model solutions to tasks as
programs,
which frames our problem as one of program induction.

@@

TODO: talk about existing works as special cases of, and points of inspiration for, \system:
\begin{itemize}
  \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
  \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
    $q$;
    % because this is a primary inspiration, might be worth further mention:
    % different program representation (routing vs. substitution; simple vs polymorphic types);
    % different grammar induction (sequitur vs. fragment grammar)
  \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
    DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
    network $q$.
    % although RobustFill has very different enumeration
\end{itemize}

@@

\textbf{Neural recognition model: tractably maxing $J$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
frontier $\mathcal{F}_x$ by searching for programs maximizing
$q(p|x)$.
The network $q$ exploits the structure of the DSL $\mathcal{D}$:
rather than directly predicting a distribution over $p$ conditioned on $x$,
it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
This approach implements an amortized
inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
Fig.~\ref{graphicalModel}.
